Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-24 tacl.tex
Responded to Tim's section 3 comments

The first step in the process is to derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this section mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is an $n$-gram frequency threshold, which for us (following \newcite{Brooke15b}) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}), excepting those highly particular proper names which we consider outside the scope of FS identification. 
% TJB: is this a subset of proper names? the "excepting ..." qualification is a bit vague
Note that our set of $n$-grams also includes gapped or non-contiguous $n$-grams. The precise definition of gaps is language specific (see \secref{sec:evaluation} for details) but the general idea is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence \cite{Wasow:2002}.
% TJB: added citation and removed "this is confirmed by empirical analysis" as it begs details of the specific analysis, which I think is more of a distraction than anything else at this point
 In a POS-tagged corpus, potential gaps can be identified by a POS regular expression. This greatly lowers the number of potentially gapped $n$-gram types, increasing precision and efficiency for negligible loss of recall. We also exclude punctuation and lemmatize the corpus.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination. 
% TJB: this last sentence is a bit opaque

Once a set of relevant $n$-grams is identified and counted, other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context, as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for word $w_i$ in the context of a word sequence $w_0, w_i, ..., w_{n-1}$ with POS tag sequence $t_0, t_1, ..., t_{n-1}$, is given by:
\begin{displaymath}
\LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}
% TJB: clarify the semantics of the $w_{j,k}$ and $t_{j,k}$ notation

%These conditional probabilities can be derived in the standard way from $n$-gram and POS $n$-gram counts. 
% TJB: removed; seems to state the obvious
We use the same equation for gapped $n$-grams, with the caveat that quantities involving sequences which include the location where the gap occurs are derived from special gapped $n$-gram statistics. Note that the identification of the best ratio across all possible choices of context, not just the largest, is important for longer FS, where the entire POS context alone might uniquely identify the phrase (and hence the word), resulting in a ratio of 1. 
% TJB: is this a good thing or bad thing? slight lack of clarity here
It also ensures that the lower bound of LPR is 1, since the ratio for a word with no context is trivially 1.

@@

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node. If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to the status of the node under consideration whose status should be considered simultaneously: since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes starting from an `all-on' state. 
% TJB: greedy how (in terms of ordering of nodes)? frequency based as per above?

In practice, we apply the following restrictions, which significantly reduce the runtime of the algorithm without sacrificing the quality of the output lexicon (based on development set testing):
\begin{itemize}
\item The complexity of the algorithm increases exponentially with the number of relevant nodes, so we limit the total number to 5. When there are more than 5 nodes turned on in the vicinity of the target node, relevant nodes are selected by ranking in terms of the influence of the nodes on each other as given by the relative change in their explainedness functions.
% TJB: say more -- relative change how?
\item To avoid having to deal with storing and processing trivial overlaps, we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering, we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}

@@

Though the results across the four corpora are reasonably similar with respect to overall f-score, there are a significant discrepancies. By using the UniDict morpheme representation as the starting unit for Japanese, the model is, in essence, doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are many more FS for Japanese. Precision is very high, and recall is relatively low. Under these conditions, methods which tend to increase precision at the cost of recall will be less desirable with respect to f-score, which explains why the full model is not the best in that one case.  Importantly, the initial $n$-gram statistics derived from corpus reflect that Japanese is different: the number of $n$-gram types, over length 4, for instance, is almost twice the number in the ICWSM, or almost 14\% of all $n$-gram types. One idea for future work is to automatically adapt to the particular properties of the input language/corpus in order to ensure a good balance between precision and recall.

>>>>

The first step in the process is to derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this section mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is an $n$-gram frequency threshold, which for us (following \newcite{Brooke15b}) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}) when proper names which we consider outside the scope of FS identification (as discussed earlier) are excluded. 
% TJB: is this a subset of proper names? the "excepting ..." qualification is a bit vague
Note that our set of $n$-grams also includes gapped or non-contiguous $n$-grams. The precise definition of gaps is language specific (see \secref{sec:evaluation} for details) but the general idea is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence \cite{Wasow:2002}.
% TJB: added citation and removed "this is confirmed by empirical analysis" as it begs details of the specific analysis, which I think is more of a distraction than anything else at this point
 In a POS-tagged corpus, potential gaps can be identified by a POS regular expression. This greatly lowers the number of potentially gapped $n$-gram types, increasing precision and efficiency for negligible loss of recall. We also exclude punctuation and lemmatize the corpus.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination: in iteration $i$, $i$-grams are filtered by the frequency threshold, and then consecutive instances of these $i$-grams are found, a set which will contain all $i+1$-grams which are also over the threshold. 
% TJB: this last sentence is a bit opaque

Once a set of relevant $n$-grams is identified and counted, other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context, as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for word $w_i$ in the context of a word sequence $w_0,..., w_i, ..., w_{n-1}$ with POS tag sequence $t_0, t_1, ..., t_{n-1}$, is given by:
\begin{displaymath}
\LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

\noindent
where $w_{j,k}$ and $t_{j,k}$ should be understood as excluding $w_{i}$ and $t_{i}$ respectively. 
% TJB: clarify the semantics of the $w_{j,k}$ and $t_{j,k}$ notation

%These conditional probabilities can be derived in the standard way from $n$-gram and POS $n$-gram counts. 
% TJB: removed; seems to state the obvious
We use the same equation for gapped $n$-grams, with the caveat that quantities involving sequences which include the location where the gap occurs are derived from special gapped $n$-gram statistics. Note that the identification of the best ratio across all possible choices of context, not just the largest, is important for longer FS, where the entire POS context alone might uniquely identify the phrase, resulting in the minimum LPR of 1 even for entirely formulaic sequences, an undesirable result. 
% TJB: is this a good thing or bad thing? slight lack of clarity here
It also ensures that the lower bound of LPR is 1, since the ratio for a word with no context is trivially 1.

@@

Given the relationships between nodes that we have defined above, it is obviously not sufficient to consider switching only the present node. If, for instance, one or more of \ex{be keep \gap under wraps}, \ex{under wraps}, or \ex{be keep \gap under} has been turned on, the covering, blocking, or overlapping effects of these other nodes will likely prevent a competing node like \ex{keep \gap under wraps} from being correctly activated. Instead, the algorithm identifies a small set of nodes (relevant nodes) which are the most important to the status of the node under consideration whose status should be considered simultaneously: since turned-off nodes have no direct effect on each other, only turned-on nodes above, below, or overlapping with the current node need be considered.  Once the relevant nodes have been identified, all nodes (including turned-off nodes) whose explainedness is affected by one or more of the relevant nodes are identified, and then a greedy search is carried out for the optimal configuration of the relevant nodes, starting from an `all-on' state and at each iteration turning off the node (if any) which most increases overall explainedness.
% TJB: greedy how (in terms of ordering of nodes)? frequency based as per above?

In practice, we apply the following restrictions, which significantly reduce the runtime without sacrificing the quality of the output lexicon (based on development set testing):
\begin{itemize}
\item We limit the total number of relevant nodes to 5. When there are more than 5 nodes turned on in the vicinity of the target node, the most relevant nodes are selected by ranking candidates by the change in explainedness across possible configurations of the target and candidate node considered in isolation.
% TJB: say more -- relative change how?
\item To avoid having to deal with storing and processing trivial overlaps, we exclude overlaps with a count of less than 5 from our lattice.
\item Many nodes have a minLPR which is a little larger than 1. There is very little chance these nodes will be activated by the algorithm, and so after applying hard covering, we do not consider activating nodes with a minLPR of less than 2.
\end{itemize}

@@

Though the results across the four corpora are reasonably similar with respect to overall f-score, there are a significant discrepancies. By using the UNIDIC morpheme representation as the starting unit for Japanese, the model is, in essence, doing an extra layer of FS identification, one which is provided by word boundaries in the other languages. The result is that there are many more FS for Japanese. Precision is very high, and recall is relatively low. Under these conditions, methods which tend to increase precision at the cost of recall will be less desirable with respect to f-score, which explains why the full model is not the best in that one case.  Importantly, the initial $n$-gram statistics derived from corpus reflect that Japanese is different: the number of $n$-gram types, over length 4, for instance, is almost twice the number in the ICWSM, or almost 14\% of all $n$-gram types. One idea for future work is to automatically adapt to the particular properties of the input language/corpus in order to ensure a good balance between precision and recall.
